# Methods {#methods}

## Experimental Design

The first data release contains data from two Sorghum growing seasons, the 4th and 6th seasons of the project:  

```{r experiments, cache = TRUE}
experiments <- tbl(bety_src, 'experiments') %>%
  dplyr::select(experiment_id = id, start_date, end_date, name, description, design)

experiments_subset <- experiments %>% 
  select(experiment_id, name, Start = start_date, Harvest = end_date, Summary = description, Design = design) %>% 
  collect() %>% 
  filter(str_detect(name,"Season 4|Season 6")) %>% 
  filter(str_detect(name, "Border Plots", negate = TRUE)) %>% 
  tidyr::separate(name, c('Season', 'Experiment'), sep = ":") 

experiments_treatments <- tbl(bety_src, 'experiments_treatments') %>%
  dplyr::select(experiment_id, treatment_id)

treatments <- tbl(bety_src, 'treatments') %>%
  dplyr::select(treatment_id = id , name, definition, control)

experiments_treatments_subset <- experiments_subset %>% 
  left_join(experiments_treatments, by = 'experiment_id', copy = TRUE) %>% 
  left_join(treatments, by = 'treatment_id', copy = TRUE)

descriptions <- list()

for(exp in 1:nrow(experiments_subset)){
  season <- experiments_subset$Season[exp]
  
  trts <- experiments_treatments_subset %>% 
    filter(Season == season) %>% 
    transmute(paste0("* *", name, "*: ", definition)) %>% unlist() %>% paste(collapse = "\n")

  t <- ifelse(experiments_treatments_subset %>% 
    filter(Season == season) %>% nrow == 1, "Treatment: ", "Treatments: ")

  out <- experiments_subset %>% 
    filter(Season == season) %>% 
    transmute( paste("\n#### ", Season, "\n\n",
                   "* Start:", Start, "\n",
                   "* End:", Harvest,"\n\n",
                   t, "\n\n",
                   Summary, "\n\n",
                   "Experimental Design: ", Design, "\n\n")) %>% 
    as.character()

  descriptions[[season]] <- paste(out, paste(t, trts), collapse = "\n")
}

writeLines(unlist(descriptions), con = '../metadata/experiments.md')
#writeLines(unlist(descriptions))
```


```{r treatments}

managements_treatments <- tbl(bety_src, 'managements_treatments') %>% 
  dplyr::select(treatment_id, management_id) 

managements <- tbl(bety_src, 'managements') %>%
  dplyr::select(management_id = id, date, mgmttype, level, units, notes) 

treatments <- tbl(bety_src, 'treatments') %>%
  dplyr::select(treatment_id = id , name, definition, control)

experiments_treatments <- tbl(bety_src, 'experiments_treatments') %>%
  dplyr::select(experiment_id, treatment_id)

managementsplus <- managements %>% 
  left_join(managements_treatments, by = 'management_id') %>%
  left_join(treatments, by = 'treatment_id') %>% 
  left_join(experiments_treatments, by = 'treatment_id') %>% 
  inner_join(experiments_subset, by = 'experiment_id', copy = TRUE) %>% 
  mutate(experiment = name) %>% 
  select(Season, date, treatment = name, activity = mgmttype, level, units, notes) %>% 
  collect()

file.remove('../book/treatments.md')
for(t in unique(managementsplus$treatment)){
  s <- managementsplus %>% 
    filter(treatment == t) %>% 
    select(Season) %>% 
    unique() %>% 
    as.character()
  tt <- managementsplus %>% 
    filter(treatment == t) %>% 
    arrange(date) %>% 
    select(-treatment, -Season) %>% 
    replace(., is.na(.), "") %>% 
    knitr::kable(format = 'markdown')
  write(paste("\n\n### ", t, "\n\n"), 
        file = '../book/treatments.md', append = TRUE)
  write(tt, 
        file = '../book/treatments.md', append = TRUE)
}

```

## Germplasm

Here we use the terms 'germplasm', 'cultivars', and 'accessions' to refer to the distinct plant genotypes that are the key unit of experimental investigations in breeding trials.

```{r germplasm}
library(jsonlite)
# Using traits package
#season_6 <-  betydb_query(table = 'experiments', name = 'MAC Season 6: Sorghum BAP')
#season_4 <-  betydb_query(table = 'experiments', name = 'MAC Season 4: All BAP Accessions')

# experiments_subset generated above
season_6 <- experiments_subset %>% 
  filter(Season == 'MAC Season 6')
season_4 <- experiments_subset %>% 
  filter(Season == 'MAC Season 4') 

# query germplasm from brapi endpoint
if(!file.exists('../metadata/season_4_germplasm.json')){
  download.file(
    paste0("https://terraref.org/brapi/v1/studies/",
           season_4$experiment_id,"/germplasm"), 
    destfile = '../metadata/season_4_germplasm.json')
}
if(!file.exists('../metadata/season_4_germplasm.json')){
  download.file(
    paste0("https://terraref.org/brapi/v1/studies/",
           season_6$experiment_id,"/germplasm"), 
    destfile = '../metadata/season_6_germplasm.json')
}
s4_germplasm <- fromJSON('../metadata/season_4_germplasm.json', flatten = TRUE)$result$data %>%
  select(germplasmName, germplasmPUI) %>% 
  mutate(season_4 = TRUE)
s6_germplasm <- fromJSON('../metadata/season_6_germplasm.json', flatten = TRUE)$result$data %>%
  select(germplasmName, germplasmPUI) %>% 
  mutate(season_6 = TRUE)
germplasm_list <- s4_germplasm %>% 
  dplyr::full_join(s6_germplasm, by = c('germplasmName', 'germplasmPUI')) 

readr::write_csv(germplasm_list, '../metadata/germplasm.csv')

germplasm_table <- germplasm_list %>% 
  knitr::kable()

germplasm_text <- c("# Germplasm \n\n A list of cultivars and the experiments in which they are planted are listed in the table below, from the file 'germplasm.csv'. Additional attributes are listed in the files 'season_4_germplasm.json' and 'season_6_germplasm.json'.\n\n", 
                        germplasm_table)

writeLines(germplasm_text, '../book/germplasm.md')
# writeLines(germplasm_text) # in appendix
```

## List of Phenotypes and How They Were Measured

Some were measured by sensor, and others were measured by hand.

```{r methods}
# Using traits package
#variables <- betydb_query(table = 'variables', limit = 'none')
#methods <- betydb_query(table = 'methods', limit = 'none')

variables <- tbl(bety_src, 'variables') 
methods <- tbl(bety_src, 'methods')

# could generalize follwoing to a function:
# get_season_data <- function(season){}

s4 <- list()
for (f in dir("../data/season_4_traits/", full.names = TRUE, pattern = 'csv')){
  s4[[f]] <- readr::read_csv(f)
}

s4 <- dplyr::bind_rows(s4) %>% 
  mutate(season = 'Season 4')

s6 <- list()
for (f in dir("../data/season_6_traits/", full.names = TRUE, pattern = 'csv')){
  s6[[f]] <- readr::read_csv(f)
}

s6 <- dplyr::bind_rows(s6) %>% 
  mutate(season = 'Season 6')

s46 <- dplyr::bind_rows(s4, s6)

s46_subset <- s46 %>% 
  dplyr::select(plot, scientificname, genotype, season, treatment, date, trait, method, mean, checked, notes, author) %>% 
  dplyr::mutate(method_type = ifelse(method %in% sensor_methods, 'sensor', 'manual'))

readr::write_csv(
  s46_subset %>% filter(season == "Season 4"), 
  '../data/season_4_traits.csv')

readr::write_csv(
  s46_subset %>% filter(season == "Season 6"), 
  '../data/season_6_traits.csv')
```

```{r s46methods}
trait_counts <- s46_subset %>% 
  group_by(trait, method) %>% 
  tally()
readr::write_csv(trait_counts, '../metadata/trait_counts.csv')

unique_traits <- unique(s46_subset$trait)
v <- variables %>% 
  select(label, name, description, units, notes)  %>% 
  filter(name %in% unique_traits) %>% 
  collect()

# alternative with R traits package:
# c <- betydb_query(table = 'citations', limit = 'none') %>% 
c <- tbl(bety_src, 'citations') %>% 
  select(author, year, title, journal, citation_id = id)

unique_methods <- unique(s46_subset$method)
m <- methods %>%  
  left_join(c, by = 'citation_id') %>% 
  filter(name %in% unique_methods) %>%
  select(name, description, author, title, year, journal) %>% 
  collect()

readr::write_csv(m, '../metadata/methods.csv')
readr::write_csv(v, '../metadata/variables.csv')

```


```{r variables.md}
variables_text <- v %>% 
#  tidyr::replace_na(list(author = "", year = "", title = "", journal = "")) %>% 
  rowwise() %>% 
  transmute(paste0("\n**", label, 
                   "** (",units,")",
                   "\n* name: ", name,
                   "\n* description: ",
                   description, 
                   "\n*_notes_: ", notes))

variables_text <- unlist(variables_text)
writeLines(text = c("Variables (Traits and Phenotypes)", 
                    variables_text),
           con  = 'variables.md')

#writeLines(variables_text)
```


```{r methods.md}
methods_text <- m %>% 
  tidyr::replace_na(list(author = "", year = "", title = "", journal = "")) %>% 
  rowwise() %>% 
  transmute(paste(paste0("\n**", name, "**"),
                  "\n* citation: ", author, year, title, journal,
                  "\n* total observations: ", trait_counts %>% filter(method == m$name[[5]]) %>% ungroup() %>% summarise(sum(n)),                   "\n* description: ", description,
                  "\n* associated variable(s):", trait_counts %>% filter(method == name) %>% select(trait) %>% paste(sep = ','), 
                  "\n\n"
                  ))

methods_text <- unlist(methods_text)
writeLines(text = c("### Methods: \n\n", methods_text), 
                    con = 'methods.md')
#writeLines(methods_text)
```

