## Phenomics

### Data

#### Raw Phenotype data

Tables of phenotypes can be found in the folders `data/raw/season_4_traits/` and `data/raw/season_6_traits/`. These directories contain one csv file for each combination of trait and measurement method. The names of these csv files help identify the contents because they follow the pattern  `[season][trait][measurement_type].csv`. For example, the file `season_6_aboveground_biomass_manual.csv` contains manual measurements of aboveground biomass taken during Season 6.

These csv files have one measurement per row for a specific date, location, genotype, and measurement. The first line contains the names of the fields:

* plot
* lat
* lon
* scientificname
* genotype
* treatment
* date: (char)
* time: provided for sensor derived data
* year: (integer): year of observation
* month: (integer): month of observation
* trait: (text) name of the trait measured. Defined in the file `metadata/variables.csv`
* method: (text) the method used to measure the trait. Defined in the file `metadata/methods.csv` 
* mean: (numeric) value of the phenotype data
* checked: (boolean) 0 = unchecked and 1 = checked
* notes: any additional information
* author: (text) name of scientist who collected the data or who wrote the algorithm used to derive phenotypes from sensor data
* season: (integer) season 4 or season 6
* method_type: (boolean) manual or sensor derived

```{r}
library(dplyr)

s4 <- list()
for (f in dir("data/raw/season_4_traits/", full.names = TRUE, pattern = 'csv')){
  s4[[f]] <- readr::read_csv(f)
}

s4 <- dplyr::bind_rows(s4) %>% 
  mutate(season = 'Season 4')

s6 <- list()
for (f in dir("data/raw/season_6_traits/", full.names = TRUE, pattern = 'csv')){
  s6[[f]] <- readr::read_csv(f)
}

s6 <- dplyr::bind_rows(s6) %>% 
  mutate(season = 'Season 6')

s46 <- dplyr::bind_rows(s4, s6)

s46_subset <- s46 %>% 
  dplyr::select(plot, scientificname, genotype, season, treatment, date, trait, method, method_type, mean, checked, author) 

trait_counts <- s46_subset %>% 
  group_by(season, trait, method_type) %>% 
  tally()

```


#### Trait Counts By Season

The number of trait records collected in each season by different methods is shown below:

**Season 4**

```{r s4_counts, include=TRUE}
trait_counts %>% 
  filter(season == "Season 4") %>% 
  ungroup() %>% 
  select(-season) %>% 
  knitr::kable(format = 'markdown',  
               format.args = list(big.mark = ","),
               caption = 'Number of observations collected in Season 4') 
```

**Season 6**

```{r s6_counts, include=TRUE}
trait_counts %>% 
  filter(season == "Season 6") %>% 
  ungroup() %>% 
  select(-season) %>% 
  knitr::kable(format = 'markdown', 
               format.args = list(big.mark = ","),
               caption = 'Number of observations collected in Season 4') 
```

### Code

For the purposes of reproducibility, the code used to run and query data from a copy of the TERRA REF trait database (terraref.org/bety) is provided. This uses Docker to run the database and R to query the data and prepare the csv files. The `code/` directory contains the following:

* R script `s4s6_sql2csv.R` that connects to the database and generates the phenomics data files described above.
* `betydb_docker/` folder that contains files named `Docker` and `docker-compose.yml` that allow a user to run the PostgreSQL database following instructions in the `README.md` file. This can be queried using the included script or from any programming language using the credentials stored in the file `.pgpass`.

